---
title: "Paired samples t-test"
output: 
  html_document:
    keep_md: yes
---

The code below simulates the running speed of someone before and after drinking x ammount of coffee.

```{r setup, include=TRUE}
names <- c()
for(i in 1:10){
  name <- paste('participant', i)
  names[i] <- name
}
before_coffee = seq(from = 5, to = 50, by = 5) + rnorm(n = 10, mean = 1, sd = 2)
coffee_sim <- data.frame( "Participants" = c(names),
                          "before_coffee" = before_coffee,
                          "after_coffee" = before_coffee + rnorm(n = 10, mean = 20, sd = 1)
)
```

Imagine you are a researcher and you want to measured the speed 
of a pearson running before and after a cup of coffee. Here's your data set. Take a look at it
compare the differences in both within subject (row to row), 
and within conditions (column to column):

```{r 1, include=TRUE}
coffee_sim
```
2) Calculate the descriptive statistics that we learned in our previous exercises

```{r 2, include=TRUE}

mean(coffee_sim$before_coffee)
mean(coffee_sim$after_coffee)
sd(coffee_sim$before_coffee)
sd(coffee_sim$after_coffee)

```
3) Using only the descriptive statistics from question 2, 
#do you think that there will be a significance difference
#in speed before and after people take coffe? Justify your answer.


4) Calculate the difference in scores between conditions:

```{r t, include=TRUE}
dif <- mean(coffee_sim$before_coffee - coffee_sim$after_coffee)
```

#5) Calculate the standard error of the mean of the differences

```{r 3, echo=TRUE}
se_mean <- sd(coffee_sim$before_coffee - coffee_sim$after_coffee)/sqrt(length(coffee_sim$before_coffee))
```

#6) Now divide the difference you found between samples with the S.E.

```{r 4, echo=TRUE}
dif/se_mean
```

#7) Now compute a paired samples t-test using r's native function. Compare the t-statistic with the result from question 6
```{r 5, echo=TRUE}
t.test(coffee_sim$after_coffee,
       coffee_sim$before_coffee,
       paired = TRUE)[["statistic"]][["t"]]

```
